---
title: "p8105_hw1_qz2527"
author: "kindle zhang"
date: "2023-09-16"
output: html_document
---

## Problem 1

```{r setup, include = FALSE}
library(tidyverse)
library(moderndive)
```

1. load the data set

```{r load data}
data("early_january_weather")
```

2. describe the data

```{r describe data}
summary(early_january_weather)
# or use head(early_january_weather)
data_row = nrow(early_january_weather)
data_col = ncol(early_january_weather)
mean_temp = mean(early_january_weather$temp)
```

the variables in this data set are: `r ls(early_january_weather)`

the size of the data set is `r data_row` rows and `r data_col` columns

the mean temputure is `r mean_temp`

3. plot 

make a data table first and use `ggplot` to plot the image

```{r plot scatter}
my_plot = tibble(
  x = early_january_weather$time_hour,
  y = early_january_weather$temp,
  humid = early_january_weather$humid
)
ggplot(my_plot, aes(x = x, y = y, group = humid)) + geom_point(aes(color = humid, size = 5))

ggsave("hw1_plot_1.png")
```
## Problem 2

Create a data frame comprised of:

a random sample of size 10 from a standard Normal distribution
a logical vector indicating whether elements of the sample are greater than 0
a character vector of length 10
a factor vector of length 10, with 3 different factor “levels”
Try to take the mean of each variable in your dataframe. What works and what doesn’t?

```{r}
my_data_2 = tibble(
  norm_data = rnorm(10),
  log_vec = norm_data > 0,
  cha_vec = c("12","22","24","123","22424","2","2424","22","24","244442"),
  fac_vec = factor(c("M","XL","L","M","XL","L","M","XL","L","M")) 
)

mean_nd = mean(pull(my_data_2, norm_data))
mean_lv = mean(pull(my_data_2, log_vec))
mean_cv = mean(pull(my_data_2, cha_vec))
mean_fv = mean(pull(my_data_2, fac_vec))
```

* we can find the mean of numeric and logical variable exists,which are `r mean_nd` and `r mean_lv` but the mean of factor and character don't exist.

the mean of the logical variable is the *proportion* of "True" in total samples

```{r}
mean_lv_2 = mean(as.numeric(my_data_2$log_vec))
mean_cv_2 = mean(as.numeric(my_data_2$cha_vec))
mean_fv_2 = mean(as.numeric(my_data_2$fac_vec))
```

we can find the character becomes numeric variable and if they are absolutely numbers, they can be calculated.

the "true" becomes 1 and "false" becomes 0, then it can be calculated.

the different levels in factor sequence becomes a sequence with a start number 1.

according to what we find above, we can explain why a mean can be calculated when we use a `as.numeric` function.
